---
title: "Insufficient sleep ~ Noise exposure "
date: today

format: html

execute: 
  echo: false
  warning: false

embed-resources: true
---

```{r}
# /////////////////////////////////////////////////////////////////////////////
# Carlos Rodriguez, PhD. CU Anschutz Dept. of Family Medicine
#
# /////////////////////////////////////////////////////////////////////////////
```

```{r}
library(tidyverse)
library(ggpattern)
library(gtsummary)
library(broom)
library(dfmtbx)
# library(splines)
library(sf)
library(spatialreg)
```

```{r}
# Run noise.qmd
# specify, but do not run census.R
# run places.R (will source census.R within)
# run models
```

```{r}
# Load data
# merged places comes from places.R script
# Used .rds to preserve geometry column. Saving as .csv removes the <sf> object functionality
data <- readRDS("D:\\rwjf_noise\\data\\prepped_data\\merged_places_2026-06-22.rds") 
```

# Number of tracts per data set
```{r}
# These were counted in the census_eda.R script.
data.frame(
  places = 2324,
  census = 2498,
  noise = 2493, # 2 are removed for santa catalina and san clemente islands,
  places_and_noise = 2000) %>% 
  format_gt_tbl()
```

# Plot of Noise exposure and Insufficient Sleep Prevalence
```{r}
data %>%
  select(pw_exp_db, SLEEP_CrudePrev) %>%
  drop_na() %>%
  ggplot(aes(x = pw_exp_db, y = SLEEP_CrudePrev)) +
  geom_point(alpha = 1) +
  geom_smooth(method = "lm") +
  theme_minimal() +
  xlab("Population Weighted Noise Exposure (dB)") +
  ylab("Insufficient Sleep Prevalence (%)")

# ggsave("sleep_vs_noise_a03.pdf")
```

<!-- # Plot of average sleep and average noise after binning into 10 noise groups -->
```{r}
# Bin the data to see if the hockey stick shape of the plot of raw values is 
# an artifact.
# binned <- data %>%
#   drop_na(pw_exp_db) %>%
#   mutate(bin = cut_number(pw_exp_db, 10)) %>%
#   group_by(bin) %>%
#   summarise(
#     mean_exposure = mean(pw_exp_db, na.rm = TRUE),
#     mean_sleep = mean(SLEEP_CrudePrev, na.rm = TRUE),
#     n = n()
#   )

# ggplot(binned, aes(mean_exposure, mean_sleep)) +
#   geom_point(aes(size = n)) +
#   geom_line() +
#   theme_minimal()
```

# Models with a dichotomous predictor of noise exposure
```{r}
# Perform Moran's I test
vars <- c("gte40db", "gte50db", "gte60db")

moran_test <- function(x) {
  # Create a copy of the main data set in the workspace
  mod_data <- data

  # Create a formula based on the input predictor variable
  formula <- as.formula(str_c("SLEEP_CrudePrev ~ ", x))
  # print(formula)

  # Fit the model
  model <- lm(formula, data = mod_data, na.action = na.exclude)

  # Get the residuals in a test data frame
  mod_data <- mod_data %>%
    mutate(residuals = residuals(model))

  # Drop the rows where values for the model are missing
  mod_data <- mod_data %>%
    filter(!is.na(residuals))

  # Plot the 2000 tracts that were included in the analysis
  centroids <- mod_data %>%
    filter(!!sym(x) == 1) %>%
    st_centroid()

  plt <- mod_data %>%
    ggplot() +
    geom_sf(aes(fill = SLEEP_CrudePrev), color = NA) +
    geom_sf(
      data = centroids,
      shape = 21,
      size = .45,
      stroke = .25,
      fill = NA,
      color = "white") +
    scale_fill_viridis_c() +
    theme_minimal()

  # Construct a neighbors list from polygon list
  # *** produced a warning about some observations not having any neighbors 
  # because the intersection of tracts with noise AND tracts with a CDC Places
  # outcome results in some tracts without any neighbors
  nb <- spdep::poly2nb(
    mod_data, 
    queen = TRUE)

  # Create the spatial weights from a neighbors list. The zero policy sets
  # weights vectors of zero length for regions without any neighbors
  lw <- spdep::nb2listw(
    nb, 
    style = "W", # Nick's team used B in this step 
    zero.policy = TRUE)

  # Moran's I test revealed p less than .05 for all dichotomous variables
  # There is evidence for spatial auto correlation, need a spatial model
  moran_results <- spdep::moran.test(
    mod_data$residuals, #cannot have missing values
    lw 
  )

  # print(moran_results)

  # Spatial error model
  sem <- errorsarlm(
    formula,
    data = mod_data,
    listw = lw,
    zero.policy = TRUE,
    na.action = na.exclude
  )

  return(list(model, plt, moran_results, sem))

}

results <- map(vars, moran_test)

```


## Unadjusted Models
## dB Greater than or equal to 40 dB
### Unadjusted regression output of a simple linear regression model
```{r}
tbl_regression(results[[1]][[1]])
```

### Plot of Insufficient Sleep Prevalence overlaid with tracts where the average dB value is greater than or equal to 40 dB
```{r}
results[[1]][[2]]
```

### Moran's I
```{r}
results[[1]][[3]]
```

### Spatial error model output
```{r}
summary(results[[1]][[4]])
```

## dB Greater than or equal to 50 dB
### Unadjusted regression output of a simple linear regression model
```{r}
tbl_regression(results[[2]][[1]])
```

### Plot of Insufficient Sleep Prevalence overlaid with tracts where the average dB value is greater than or equal to 50 dB
```{r}
results[[2]][[2]]
```

### Moran's I
```{r}
results[[2]][[3]]
```

### Spatial error model output
```{r}
summary(results[[2]][[4]])
```

## dB Greater than or equal to 60 dB
### Unadjusted regression output of a simple linear regression model
```{r}
tbl_regression(results[[3]][[1]])
```

### Plot of Insufficient Sleep Prevalence overlaid with tracts where the average dB value is greater than or equal to 60 dB
```{r}
results[[3]][[2]]
```

### Moran's I
```{r}
results[[3]][[3]]
```

### Spatial error model output
```{r}
summary(results[[3]][[4]])
```

<!-- 

### Adjusted for age, race, and ethnicity
```{r}
# Nick's lab had performed some spatial error models
# Formula 1 "y ~ median_value   + pct_homeowners  + pct_foreign_born +
#            median_income + black_count +
#            native_count + latino_count + white_count"

# formula2 <-  "y_night ~ median_value   + pct_homeowners  + pct_foreign_born +
#            median_income + black_count +
#            native_count + latino_count + white_count"
```

```{r}
# Collect the covariates for using as.formula()
demographics <- c("median_age", "prop_latino", "prop_white", "prop_black", "prop_native", "prop_asian", "prop_foreign_born")
formula <- as.formula(
  str_c(
    "SLEEP_CrudePrev ~ ",
    "gte40db + ",
    paste0(demographics, collapse = " + "))
    )
```

```{r}
mod_40b <- lm(formula, data = data)

summary(mod_40b)

car::vif(mod_40b)

plot(mod_40b)
```


### Adjusted for SES

```{r}
ses <- c("median_value", "median_income", "prop_owner_occ", "prop_renter_occ", 
  "prop_bach", "prop_masters", "prop_prof", "prop_tot_edu", 
  
  "median_rent", "median_year_built_srs")

formula <- as.formula(
  str_c(
    "SLEEP_CrudePrev ~ ",
    "gte40db + ",
    paste0(c(demographics, ses), collapse = " + "))
    )
```

```{r}
mod_40c <- lm(formula, data = data)

summary(mod_40c)
``` -->

<!-- ### Adjusted for comorbidities -->

<!-- ```{r}

```

## dB Greater than or equal to 50 dB
```{r}
mod_50a <- lm(SLEEP_CrudePrev ~ gte50db, data = data)

summary(mod_50a)
```


## dB Greater than or equal to 60 dB
```{r}
mod_60a <- lm(SLEEP_CrudePrev ~ gte60db, data = data)

summary(mod_60a)
```
 -->


<!-- ////////////////////////////////////////////////////////////////////// -->
<!-- 
# Linear model
```{r}
mod_lm <- lm(SLEEP_CrudePrev ~ pw_exposure, data = data)

summary(mod_lm)

plot(mod_lm)
```

```{r}
#| echo: false
#| eval: false

# Interpretation of linear model
# The coefficient for pw_exposure is 0.22. This is interpreted as a 1-unit 
# increase in noise exposure is associated with a .22 unit increase in 
# insufficient sleep. Small but significant effect
```


# Polynomial model with a quadratic term
```{r}
# Center the noise exposure variable
data <- data %>%
  mutate(pw_exposure_c = pw_exposure - mean(data$pw_exposure, na.rm = TRUE))

mod_pol <- lm(
  SLEEP_CrudePrev ~ pw_exposure_c + I(pw_exposure_c^2),
  data = data
)

summary(mod_pol)

plot(mod_pol)
```

```{r}
#| eval: false
#| echo: false

# Interpretation of polynomial model coefficients using the centered version 
# of pw_exposure
# 1. pw_exposure. The population weighted exposure coefficient is .277 and is 
# significant. The slope between noise and insufficient sleep is .28 at '
# the average level of noise.
# 2. the quadratic term is 0.008. It's positive so that for each unit increase
# in x, the slope increase by 2 * 0.008, so the effect of noise is magnified
# as it increases.

```

## Test of the quadratic term in the polynomial model vs the linear model
```{r}
model_lm <- lm(
  SLEEP_CrudePrev ~ pw_exposure_c,
  data = data)
anova(mod_lm, mod_pol)
```

# Piecewise linear model with the inflection point set to 50db
```{r}
data <- data %>%
  mutate(
    above50 = ifelse(pw_exposure > 50, pw_exposure - 50, 0)
  )

mod_pw <- lm(
  SLEEP_CrudePrev ~ pw_exposure + above50,
  data = data
)

summary(mod_pw)

plot(mod_pw)
```

```{r}
#| eval: false
#| echo: false

# Gives two slopes, one before and after the inflection pointat 50dB
```

# Splines model
```{r}
knots <- quantile(
  data$pw_exposure,
  probs = c(0.2, 0.4, 0.6, 0.8),
  na.rm = TRUE
)

mod_spl <- lm(
  SLEEP_CrudePrev ~ ns(
    pw_exposure,
    knots = knots),
  data = data
)
```

```{r}
summary(mod_spl)

plot(mod_spl)
```

# Compare AICs
```{r}
AIC(mod_lm, mod_pol, mod_pw, mod_spl)
```

 -->
