---
title: "RWJF Helicopter Noise - Data Processing Workflow"
format: 
  html:
    toc: true

execute:
  warning: false
  message: false # Also hides package loading messages

embed-resources: true
---


```{r}
#| eval: false
#| echo: false
# -----------------------------------------------------------------------------
# Carlos Rodriguez, PhD. CU Anschutz Dept. of Family Medicine
# RWJF HelicopterSleep

# Data processing
#  Tract & county maps are from year = 2020
#  Noise is from 2020 annual
#  Population data is from
# -----------------------------------------------------------------------------
```

```{r, warning = FALSE, message = FALSE}
#%% Libraries
  library(tidyverse)
  library(tidycensus)
  library(sf)
  library(tidyterra)
  library(terra)
  library(exactextractr)
```


# Load Landscan data
```{r}
#%% Load the landscan population raster stored locally
# has resolution of 1/1200 degree
# coordinate refeference system (crs) is 4326
pop <- rast("D:\\rwjf_noise\\data\\landscan\\landscan-mosaic-unitedstates-v1-assets\\landscan-mosaic-unitedstates-v1.tif")

# crs(pop)
# res(pop)
```

# Load UCIC noise data (currently using Annual Average from 2020)
```{r}
#%% Load and prepare noise data 
# File is .csv format with a noise level for each lat and long coordinates.
noise_file <- "D:\\rwjf_noise\\data\\noise\\annual\\Noise_Annual_2020.csv"

# Rename coordinate columns for convenience.
exposure_df <-read_csv(noise_file, show_col_types = FALSE) %>%
    rename(lon = Longitude, lat = Latitude) %>%
    mutate(L_night_ge32 = ifelse(L_night >= 32, 1, 0))
```

## Display data structure
```{r}
str(exposure_df)
```

## Display a sample of rows
```{r}
# Display noise exposure data frame
exposure_df %>% 
  slice_sample(n = 6)
```


# Load LA County tracts
```{r}
#%% Prepare LA County tracts object
options(tigris_progress_bar = FALSE)
options(tigris_use_cache = TRUE)

suppressMessages({
# Load LA County tracts as simple features (sf) from tidycensus
# crs loads as 4269
tracts <- get_acs(
  geography = "tract",

  # Corresponds to total population variable, not used but argument needs to be
  # set for the function to work
  variables = "B01001_001E",   
  state = "CA",
  county = "Los Angeles",
  geometry = TRUE,
  year = 2020
)
})

# crs(tracts)
```

## Plot raw noise data overlaid with LA county census tracts
### Create a simple features object from raw noise data
```{r}
# Set seed for sample_frac()
set.seed(12345)

# Samples a fraction of the exposure data frame for plotting, otherwise it's too
# computationally intensive to plot. Complete data set is over 2million rows.
exposure_sf <- st_as_sf(
  exposure_df %>% 
    sample_frac(.1), 
  coords = c("lon", "lat"), 
  crs = 4326)

# Transform tracts crs from 4269 to 4326
tracts <- st_transform(tracts, 4326)

# Check if both exposure and tracts have the same crs
st_crs(exposure_sf) == st_crs(tracts)
```

### Create a clipping polygon to crop maps to excude San Celemente and Santa Catalina Islands
```{r}
# Get the minimum latitude value from the exposure data to use as 
# a cutoff for removing the islands
lat_cutoff <- min(exposure_df$lat)

# Create a bb object of LA County Tracts
bb <- st_bbox(tracts)

# Modify the bb object's y minimum value
# uses the previously defined cut off value above to raise the souther edge
bb["ymin"] <- lat_cutoff   

# Create an sfc object from the modified bb object that will be used to clip
# away the islands
clip_poly <- st_as_sfc(bb)
```

### Crop tracts
```{r}
# Crop the tracts by taking the overlap between the tracts sf object
# and the clipping polygon.
la_tracts_cropped <- st_intersection(tracts, clip_poly)
```

### Plot raw noise data
- white specks appear due to random sampling of dat points
```{r}
# Plot the data using the sf objects
ggplot() +
  geom_sf(data = exposure_sf, aes(color = L_night), size = 0.5, alpha = 0.8) +
  geom_sf(data = la_tracts_cropped, fill = NA, color = "white", alpha = 0.8, linewidth = .1) +
  coord_sf() +
  labs(
    title = "Average 2020 L_night values (dB) in LA County (not in Landscan space)",
    color = "Ldn (dB)") +
  theme_minimal()
```

# Prepare LA County tracts
Workflow of tracts is from sf --> spatial vector
```{r}
# remove rows where geometry is empty
# Tracts where GEOID end in 9901:9903 with no geometry are removed
# Empty tracts must be removed to successfully perform exact_extract() in a
# downstream step, otherwise "fail to get geometry extent" errors arise.
# la_tracts_cropped <- la_tracts_cropped %>%
#   filter(!st_is_empty(la_tracts_cropped))

# Convert to spatial vector to be used for cropping
la_vect <- vect(la_tracts_cropped)

# Check if the tracts and pop have the same crs
same.crs(la_vect, pop)
```

# Crop Landscan to pixels within the LA county
```{r}
#%% Crop the landscan population raster
# crop and mask have different functions, crop is to establish the bounding box
# of pop to that of la_vect, and mask is to set any value outside the tracts, 
# but within the bbox to NA
la_pop <- mask(crop(pop, la_vect), la_vect)
```

```{r}
ggplot() +
  geom_spatraster(data = la_pop) +
  geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
  labs(
    title = "Landscan pop values",
    fill = "") +
  theme_minimal()
```

# Previous Workflow for transforming noise exposure data to raster for population weighting
## Convert exposure data frame to spatial vector
Previous worfklow was exposure data frame --> spatial vector --> raster in landscan space. Resulted in multiple points getting assigned to landscan pixels requiring a need to average out pixels before weighting.
```{r}
# Convert exposure_df to terra spatial vector (SpatVector) with point geometry
exposure_vect <- vect(
  exposure_df,
  geom = c("lon", "lat"),
  crs = "EPSG:4326"
)

# Check that the exposure vect is in the same coordinate reference system
same.crs(la_pop, exposure_vect)
```

## Rasterize exposure vector using the length() function
This step counts the number of noise data points that fall within a single landscan pixel
```{r}
#%% Get a count of noise measurements per pixel
exp_count <- rasterize(
  exposure_vect,
  la_pop,
  fun = length,
  background = 0
)
```

## Descriptives for exp_count raster
```{r}
# Check the max number of points assigned to a cell.
# The max value is 2
global(exp_count, "max", na.rm = TRUE)
```

```{r}
# Check the number of pixels where the number of points per pixel is greater
# than 1.
# Results in 383,180 cells with exactly 2 points per pixel.
global(exp_count > 1, "sum", na.rm = TRUE)
```

```{r}
# Total number of pixels
# 1,599,360 in LA county
sum(values(exp_count) > 0, na.rm = TRUE)
```

```{r}
# Proportion of cells that have 2 points (noise values) per pixel
# is approximately 24%
383180/1599360
```

```{r}
# Tabulation of the frequencies of the values in exp_count pixels
# 2180348 pixels have values of 0, 1216180 have values of 1, 383180 have values of 2
freq(exp_count)
```

## Plot exp_count raster
Shows a plaid like pattern of how noise points are assigned to pixels resulting in more than one noise values per pixel and requiring an aggregating function to achieve 1 noise value per pixel.
```{r}
# Plot exposure counts (number of points per pixel)
ggplot() +
    geom_spatraster(data = exp_count) +
    geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
    labs(
      title = "Exposure points per land scan pixel") +
  theme_minimal()
```

## Data QA on exposure_df
```{r}
#| eval: false
# Inspect the spacing of the exposure df
head(sort(unique(exposure_df$lon)))
head(sort(unique(exposure_df$lat)))

# Spacing of each lon point is 0.001091187
diff(sort(unique(exposure_df$lon)))[1:10]

# Spacing of each lat point is 0.0006719639 
diff(sort(unique(exposure_df$lat)))[1:10]

# Check resolution of lands scan population raster
# 0.0008333333 x 0.0008333333
res(pop)


# Exposure_df is rectangular shaped, lon is larger than lat
# whereas landscan population is square
```

```{r}
#| eval: false
# There are 6 unique values for the spacing in lon, but they all display as 
# 0.001091187 degrees possibly due to floating point precision in rounding.
length(unique(diff(sort(unique(exposure_df$lon)))))

# There are 2 unique values for the spacing in lat, but they all display as 
# 0.0006719639 degrees
length(unique(diff(sort(unique(exposure_df$lat)))))

# If values are rounded to 8 places after the decimal, only 1 unique value for
# the spacing in each dimension
sort(unique(round(diff(sort(unique(exposure_df$lon))), 8)))

sort(unique(round(diff(sort(unique(exposure_df$lat))), 8)))

length(unique(exposure_df$lon))
length(unique(exposure_df$lat))
```

# Updated workflow
Since the exposure data is already in a regularly spaced grid, another approach is to rasterize the noise exposure data to its native grid after importing from .csv file and then resampling to the landscan grid.

## Rasterize exposure data frame
* Perform a linear transformation of the dB values before rasterizing
```{r}
# Linearize the dB values
exposure_df <- exposure_df %>%
  mutate(L_night_lin = 10 ^ (L_night/10))

# Create raster directly from points, instead of points to spatial vector to raster
exposure_rast <- rast(
  exposure_df[, c("lon", "lat", "L_night_lin")],
  type = "xyz",
  crs = "EPSG:4326"
)
```

```{r}
# Approximately 111.1 meeters by 66.7 meters.
res(exposure_rast)
```

## Resample the exposure raster
* Resample the linear transformed variables
```{r}
# Resample the exposure raster to landscan grid
exp_landscan_lin <- resample(
  exposure_rast,
  la_pop,
  method = "bilinear"
)
```

## Transform back to dB
```{r}
# Convert exposure raster back to dB
exp_landscan_db <- 10 * log10(exp_landscan_lin)
```

```{r}
# Check the resolutions of exp_landscan and la_pop
res(exp_landscan_lin)
res(exp_landscan_db)
res(la_pop)
```

```{r}
#| eval: false 
#| echo: false


## Plot the cropped exposure raster overlaid with tracts

ggplot() +
    geom_spatraster(data = exp_landscan_db) +
      geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
        labs(
            title = "Average 2020 L_night values (dB) in LA County (resampled to Landscan space)",
                fill = "dB") +
                  theme_minimal()

ggplot() +
  geom_spatraster(data = la_pop)

```

# Approach 1: Calculate the tract level population-weighted values of noise exposure using the linear transformed dB values
```{r}
# Compare la_pop and exp_landscan geometries
# At this point la_pop is cropped to exclude Santa Catalina and San Celemente islands
compareGeom(la_pop, exp_landscan_lin)

# is pixel valid, is not missing 
# Creates another raster of logicals where a population value is not missing
valid <- !is.na(la_pop)
```

```{r}
#| eval: false
#| echo: false

# ggplot() +
#   geom_spatraster(data = valid) +
#   geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
#   labs(
#     Vitle = "Valid population exposure pixels", 
#     # fill = "dB"
#     ) +
#   theme_minimal()
```

```{r}
#| eval: false
#| echo: false
# ggplot() +
#   geom_spatraster(data = la_pop) +
#   geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
#   labs(
#     Vitle = "Valid population exposure pixels", 
#     fill = "dB") +
#   theme_minimal()
```

```{r}
weighted <- exp_landscan_lin * la_pop
```

```{r}
minmax(weighted)
```

```{r, message = FALSE}
#%% Calculated a weighted exposure variable & plot
num <- exact_extract(weighted, la_tracts_cropped, 'sum', progress = FALSE)

den <- exact_extract(la_pop, la_tracts_cropped, 'sum', progress = FALSE)

# Set the population weighted exposure variable by dividing num by den
la_tracts_cropped$pw_exp_lin <- num / den

# Convert the linearized noise levels back to dB
la_tracts_cropped$pw_exp_db <- 10 * log10(la_tracts_cropped$pw_exp_lin)
```

## Plot population weighted exposure (dB) by tract
* Approach 1 = noise --> linear transform --> resample --> aggregate --> transform to dB
```{r}
# Plot population weighted exposure by tract
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = pw_exp_db)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```

```{r}
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = pw_exp_db2)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```


# Approach 2: Calculate the tract level population-weighted values of noise exposure after converting the resampled linear values to dB
```{r}
weighted <- exp_landscan_db * la_pop
```

```{r}
minmax(weighted)
```

```{r}
#%% Calculated a weighted exposure variable & plot
num <- exact_extract(weighted, la_tracts_cropped, 'sum', progress = FALSE)
den <- exact_extract(la_pop, la_tracts_cropped, 'sum', progress = FALSE)

# Set the population weighted exposure variable by dividing num by den
la_tracts_cropped$pw_exp_db2 <- num / den
```

## Plot population weighted exposure (dB) by tract
* Approach 1 = noise --> linear transform --> resample --> transform to dB --> aggregate
```{r}
# Plot population weighted exposure by tract
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = pw_exp_db2)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```

# Differences between the two approaches
```{r}
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(diff = pw_exp_db2 - pw_exp_db)
```

```{r}

```

```{r}
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = diff)) +
    scale_fill_viridis_c() +
    labs(title = "Differences in population weighted exposure values (db2 - db)") +
    theme_minimal()
```

```{r}
la_tracts_cropped %>%
  summarise(
    mean_diff = mean(diff),
    min_diff = min(diff),
    max_diff = max(diff)
  )
```

```{r}
# Approach 1 is noise --> linear transform --> resample --> aggregate --> transform to dB
# Approach 2 is noise --> linear transform --> resample --> transform to dB --> aggregate

# Approach 2 > approach 1?
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(db2_gt_db1 = ifelse(pw_exp_db2 > pw_exp_db, 1, 0))

# Proportion of tracts where the values from Approach 2 are greater than Approach 1
mean(la_tracts_cropped$db2_gt_db1)

# Only 5 tracts in Approach 2 are greater than Approach 1
sum(la_tracts_cropped$db2_gt_db1)

# Approach 1 > approach 2?
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(db1_gt_db2 = ifelse(pw_exp_db > pw_exp_db2, 1, 0))

# Proportion of tracts where the values from Approach 1 are greater than Approach 2
# Most tracts in Approach 1 end up with higher values than Approach 2
mean(la_tracts_cropped$db1_gt_db2)

sum(la_tracts_cropped$db1_gt_db2)
```

<!-- # Diagnostic plots -->
```{r}
#| eval: false
#| echo: false
# ggplot() +
#     geom_spatraster(data = la_pop) +
#     geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
#     labs(
#       title = "Landscan") +
#   theme_minimal()
```

```{r}
#| eval: false
#| echo: false
# minmax(la_pop)
```

<!-- ## Proportion of populated pixels within each tract -->
```{r}
#| eval: false
#| echo: false
# % of pixels in a tract that have a noise value AND have people living in it.
# Use max, because it will either be one or 0
# exposure_presence_rast <- rast(
#   exposure_df[, c("lon", "lat", "L_night_ge32")],
#   type = "xyz",
#   crs = "EPSG:4326"
# )

# exp_pres_landscan <- resample(
#   exposure_presence_rast,
#   la_pop,
#   # method = "near" # sets nearest neighbor assignment as the method
#   method = "bilinear"
# )

# # Create a new raster of LA population where the pixel value is greater than 0
# la_pop_presence <- la_pop > 0

# compareGeom(la_pop_presence, exp_pres_landscan)


# # Sum of pixels within each tract where the pixel is > 0
# num <- exact_extract(la_pop_presence, la_tracts_cropped, 'sum')

# # Sum of all pixels within a tract
# den <- exact_extract(la_pop, la_tracts_cropped, function(values, coverage_fraction) {
#   sum(coverage_fraction)
# })

# la_tracts_cropped$prop_pop_pixels <- num / den
```

```{r}
#| eval: false
#| echo: false
# Plot the proportion of populated pixels
# la_tracts_cropped %>%
#   ggplot() +
#     geom_sf(aes(fill = prop_pop_pixels)) +
#     scale_fill_viridis_c() +
#     labs(title = "Proportion of Populated Pixels") +
#     theme_minimal()
# ggsave("prop_populated_pixels.pdf")
```

<!-- Binarize at 40, 50, and 60 dB -->
```{r}
#| eval: false
#| echo: false

la_tracts_cropped %>%
  mutate(
    gt40db = ifelse(db2 > 40, 1, 0),
    gt50db = ifelse(db2 > 50, 1, 0),
    gt60db = ifesle(db2 > 60, 1, 0)
    )
```

# Export tracts data
```{r}
# Output file to .rds
# saveRDS(la_tracts_cropped, "D:\\rwjf_noise\\data\\noise\\lac_pw_noise.rds")
```

```{r}
#| eval: false
#| echo: false

#%% Custom mean_db function
# names(exposure_vect) # To display the available variable names during
# development
# L_* variables are in dB scale.
# L_bar = L - 10 * log10(n)
# L = 10 * log10(sum(10^(x/10)))
# eq.s refer to Onosokki document from Yair
mean_db <- function(x) {
  # Calculate the average power given a vector of values in decibel(dB) scale 
  # Only average non-missing values
  x <- x[!is.na(x)]

  # if there are no values within a pixel return NA
  if (length(x) == 0) return(NA_real_)
  
  # eq. 12-5
  # 10*log10(sum(10^(x/10))) - (10 * log10(length(x))) 

  #Alternative form from eq. 12-6, bc each p_n^2 / p_0^2 = 10^(L_n/10)
  10 * log10(mean(10^(x/10))) 
}

# Check the log_mean_db function
# This should return 77.4, per example equation 12-6 from Onosokki document
# x  <- c(80, 70)
# mean_db(x)
```

```{r}
#| eval: false
#| echo: false


#%%
# Rasterize() essentially assigns points to landscan cells/pixels
exposure_raster <- rasterize(
  exposure_vect,
  la_pop,
  field = "L_night",
  fun = mean_db,
  background = NA
)


#%% Crop the exposure
# exposure cropped, average dB per pixel
exp_crop <- mask(crop(exposure_raster, la_vect), la_vect)

compareGeom(la_pop, exp_crop)

# is pixel valid, is not missing 
# Creates another raster of logicals where a dB value is not missing
valid <- !is.na(exp_crop)

# Adjust to valid non-missing pixels
pop_adj <- la_pop * valid # removes cells where noise is missing
weighted <- exp_crop * pop_adj



#%% Calculated a weighted exposure variable & plot
num <- exact_extract(weighted, la_tracts, 'sum')
den <- exact_extract(pop_adj, la_tracts, 'sum')

# Set the population weighted exposure variable by dividing num by den
la_tracts$pw_exposure <- num / den


# Plot population weighted exposure by tract
la_tracts %>%
  drop_na(pw_exposure) %>%
  ggplot() +
    geom_sf(aes(fill = pw_exposure)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
  ggsave("pw_exposure_x_tract.pdf")

```

```{r}
#| eval: false
#| echo: false


#%% Diagnostic plot
# % of pixels in a tract that have a noise value AND have people living in it.

# Use max, because it will either be one or 0
exposure_presence_raster <- rasterize(
  exposure_vect,
  pop,
  field = "L_night_ge32",
  fun = max,
  background = 0
)

# Mask & crop exposure presence to tract
exp_present_crop <- mask(crop(exposure_presence_raster, tracts_vect), tracts_vect)

# Binarize the population values
pop_presence <- pop > 0

# Mask & crop population presence to tract
pop_present_crop <- mask(crop(pop_presence, tracts_vect), tracts_vect)

compareGeom(pop_present_crop, exp_present_crop)


num <- exact_extract(pop_present_crop, tracts, 'sum')
den <- exact_extract(pop_crop, tracts, function(values, coverage_fraction) {
  sum(coverage_fraction)
})

tracts$prop_pop_pixels <- num / den

# Output file to .rds
saveRDS(tracts, "D:\\rwjf_noise\\data\\noise\\lac_pw_noise.rds")

# Plot the proportion of populated pixels
tracts %>%
  drop_na(pw_exposure) %>% #used for filtering only
  ggplot() +
    geom_sf(aes(fill = prop_pop_pixels)) +
    scale_fill_viridis_c() +
    labs(title = "Proportion of Populated Pixels") +
    theme_minimal()
ggsave("prop_populated_pixels.pdf")

mean(tracts$prop_pixels_w_value)

# /////////////////////////////////////////////////////////////////////////////
# Both noise greater than 32 and population present
both <- pop_present_crop * exp_present_crop

num <- exact_extract(both, tracts, 'sum')
den <- exact_extract(pop_present_crop, tracts, 'sum')

tracts$prop_exposed_pixels <- num / den

tracts %>%
  drop_na(pw_exposure) %>% #used for filtering only
  ggplot() +
    geom_sf(aes(fill = prop_exposed_pixels)) +
    scale_fill_viridis_c() +
    labs(title = "Proportion of Populated Pixels with Exposure >= 32dB") +
    theme_minimal()
ggsave("prop_exposed_pixels_ge32.pdf")


ggplot() +
  geom_sf(
    data = exposure_sf %>% slice_sample(prop = 0.1), 
    aes(color = L_night), 
    size = 0.5) +
  geom_sf(
    data = tracts, 
    fill = NA, 
    color = "black") +
  scale_color_viridis_c() +
  theme_minimal()
```