---
title: "Insufficient sleep ~ Noise exposure "
date: today

format: 
  html:
    toc: true

execute: 
  echo: true
  warning: false

embed-resources: true
---

```{r}
#| echo: false
#| eval: false
# /////////////////////////////////////////////////////////////////////////////
# Carlos Rodriguez, PhD. CU Anschutz Dept. of Family Medicine
#
# /////////////////////////////////////////////////////////////////////////////
```

```{r}
library(tidyverse)
library(ggpattern)
library(gtsummary)
library(broom)
library(dfmtbx)
library(splines)
library(sf)
library(spatialreg)
```

```{r}
#| echo: false
#| eval: false
# Run noise.qmd
# specify, but do not run census.R
# run places.R (will source census.R within)
# run models
```

```{r}
# Load data
# merged places comes from places.R script
# Used .rds to preserve geometry column. Saving as .csv removes the <sf> object functionality
data <- readRDS("D:\\rwjf_noise\\data\\prepped_data\\merged_places_2026-06-22.rds") 
```

# Number of tracts per data set
* n.b. these represent totals. The number of observations where a CDC places outcome, census covariates, and noise exposure intersect will depend on the covariates chosen for a given model.
```{r}
# These were counted in the census_eda.R script.
data.frame(
  places = 2324,
  census = 2498,
  noise = 2493, # 2 are removed for santa catalina and san clemente islands,
  places_and_noise = 2000) %>% 
  format_gt_tbl()
```

# Comparisons of tracts with the CDC places outcome variable vs those without along dimensions of the CDC Covariates.
* Two sample t-tests for continuous variables, Pearson's Chi-Square test for binary variables. Population weighted exposure in dB was converted to linear scale for calculating mean, sd, and pvalue.
```{r}
# Create a binary indicator for having the sleep outcome variable
data <- data %>%
  mutate(
    has_cdc_sleep = ifelse(is.na(SLEEP_CrudePrev), 0, 1),
    # Using log will pose problems for those tracts with 0 counts, log1p() instead
    log_black_count = log1p(black_count),
    log_native_count = log1p(native_count),
    log_latino_count = log1p(latino_count),
    log_white_count = log1p(white_count),
    log_asian_count = log1p(asian_count)
    )
```

```{r}
has_sleep_comparisons <- data %>%
  as.data.frame() %>%
  select(
    has_cdc_sleep, 
    total_pop,
    pop_density,
    median_value, 
    pct_homeowners,
    pct_college,
    pct_foreign_born, 
    median_income, 
    black_count, 
    native_count, 
    latino_count, 
    white_count,
    asian_count,
    log_black_count, 
    log_native_count, 
    log_latino_count, 
    log_white_count,
    log_asian_count,
    gte40db,
    gte50db,
    gte60db,
    ) %>%
  tbl_summary(
    by = has_cdc_sleep,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
        )) %>%
  add_p(
    test = all_continuous() ~ "t.test",
    test.args = all_tests("t.test") ~ list(var.equal = TRUE)
  ) %>%
  as_tibble()
```

```{r}
#| echo: false
#| eval: false
View(data %>%
  as.data.frame() %>%
  select(total_pop, TotalPopulation) %>%
  pivot_longer(cols = everything(), names_to = "type", values_to = "population") %>%
  mutate(type = ifelse(type == "total_pop", "ACS", "CDC")) %>%
  tbl_summary(
    by = type,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})")) %>%
  add_p(
    test = all_continuous() ~ "t.test",
    test.args = all_tests("t.test") ~ list(var.equal = TRUE)) %>%
  as_tibble())
```

```{r}
# This chunk performs the statistical comparison of the population weighted 
# exposure variable in linearized dB as opposed to the log scale, this is 
# used for generating a p-value.
pw_exp_db_lin_results <- data %>%
  as.data.frame() %>%
  select(
    has_cdc_sleep, 
    pw_exp_lin
    ) %>%
  tbl_summary(
    by = has_cdc_sleep,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
        )) %>%
  add_p(
    test = all_continuous() ~ "t.test",
    test.args = all_tests("t.test") ~ list(var.equal = TRUE)
  ) %>%
  as_tibble()

# Comparison of the values of noise exposure.
# This is to display the row for dB scale and the p value from the linear scale
pw_exp_db_results <- data %>%
  st_drop_geometry() %>%
  group_by(has_cdc_sleep) %>%
  summarise(
    mean_db_lin = mean(pw_exp_lin),
    sd_db_lin = sd(pw_exp_lin)
  ) %>%
  mutate(across(mean_db_lin:sd_db_lin, ~ round(10 * log10(.x), 2))) %>%
  mutate(mean_sd = str_c(mean_db_lin, " (", sd_db_lin, ")")) %>%
  select(has_cdc_sleep, mean_sd) %>%
  pivot_wider(values_from = mean_sd, names_from = has_cdc_sleep) %>%
  mutate(
    `**Characteristic**` = "pw_exp_db",
    `**p-value**` = pw_exp_db_lin_results$`**p-value**`
    ) %>%
  rename(`**0**  \nN = 493` = `0`, `**1**  \nN = 2,000` = `1`) %>%
  select(3, everything())
```

```{r}
# Create a final output table.
final_output <- bind_rows(
  has_sleep_comparisons,
  pw_exp_db_results) 
  
old_names <- names(final_output)
new_names <- str_remove_all(old_names, fixed("**"))
names(final_output) <- new_names

final_output %>%
  format_gt_tbl()
```



# Plot the tracts without a CDC Insufficient Sleep variable
* white dot indicates a CDC Sleep variable is available
* total_pop captured from ACS
```{r}
centroids <- data %>%
  filter(has_cdc_sleep == 1) %>%
  st_centroid()

data %>%
  ggplot() +
  geom_sf(aes(fill = total_pop), color = NA) +
  geom_sf(
    data = centroids,
    shape = 21,
    size = .45,
    stroke = .25,
    fill = NA,
    color = "white") +
  scale_fill_viridis_c() +
  theme_minimal()
```

```{r}
#| echo: false
#| eval: false

# Determine how many tracts have 0 total_pop and has an insufficient sleep dv
# 17 total tracts have an ACS total_pop value of 0
# 2 of those tracts have a cdc sleep variable
data %>% 
  as.data.frame() %>%
  filter(total_pop == 0) %>% 
  select(has_cdc_sleep) %>%
  tbl_summary()
```

```{r}
#| echo: false
#| eval: false

# Those that have a cdc_sleep value, and have a total population value of 0
# what is the population value from the CDC places data set
data %>% 
  as.data.frame() %>%
  filter(total_pop == 0) %>% 
  select(TotalPopulation, has_cdc_sleep) %>%
  tbl_summary()
```

# Plot of Noise exposure in dB and Insufficient Sleep Prevalence
```{r}
data %>%
  select(pw_exp_db, SLEEP_CrudePrev) %>%
  drop_na() %>%
  ggplot(aes(x = pw_exp_db, y = SLEEP_CrudePrev)) +
  geom_point(alpha = 1) +
  geom_smooth() +
  theme_minimal() +
  xlab("Population Weighted Noise Exposure (dB)") +
  ylab("Insufficient Sleep Prevalence (%)")

# ggsave("sleep_vs_noise_a03.pdf")
```

# Plot of Noise exposure in linear scale and Insufficient Sleep Prevalence
```{r}
# Plot the linear version
data %>%
  select(pw_exp_lin, SLEEP_CrudePrev) %>%
  drop_na() %>%
  ggplot(aes(x = pw_exp_lin, y = SLEEP_CrudePrev)) +
  geom_point(alpha = 1) +
  geom_smooth() +
  theme_minimal() +
  xlab("Population Weighted Noise Exposure (dB)") +
  ylab("Insufficient Sleep Prevalence (%)")
```

<!-- # Plot of average sleep and average noise after binning into 10 noise groups -->
```{r}
#| echo: false
#| eval: false
# Bin the data to see if the hockey stick shape of the plot of raw values is 
# an artifact.
# binned <- data %>%
#   drop_na(pw_exp_db) %>%
#   mutate(bin = cut_number(pw_exp_db, 10)) %>%
#   group_by(bin) %>%
#   summarise(
#     mean_exposure = mean(pw_exp_db, na.rm = TRUE),
#     mean_sleep = mean(SLEEP_CrudePrev, na.rm = TRUE),
#     n = n()
#   )

# ggplot(binned, aes(mean_exposure, mean_sleep)) +
#   geom_point(aes(size = n)) +
#   geom_line() +
#   theme_minimal()
```

```{r}
# Create the additive log-ratio transformations for the race and ethnic categories
data <- data %>%
  mutate(
    black_alr = log((black_count + 1)  / (white_count + 1)),
    native_alr = log((native_count + 1) / (white_count + 1)),
    latino_alr = log((latino_count + 1) / (white_count + 1)),
    asian_alr = log((asian_count + 1)/ (white_count + 1)))
```

```{r}
# Get the latitude and longitude of the tracts
# *** Adding latitude and longitude caused errorsarlm models to error out with
# a "system is computationally singular" message.
data <- data %>%
  mutate(
    centroid = st_centroid(geometry))

coords <- st_coordinates(data$centroid)

data <- data %>%
  mutate(
    longitude = coords[, 1],
    latitude  = coords[, 2]) %>%
  select(-centroid)


# Re-scale the lat and long coordinates
# re-scaling prevents "system is computationally singular" error
data <- data %>%
  mutate(
    lon_z = scale(longitude),
    lat_z = scale(latitude))
```

```{r}
# Create the distance from downtown LA variable
tracts_utm <- st_transform(data, 26911)

dtla <- st_sfc(
  st_point(c(-118.2437, 34.0537)),
  crs = 4326
)

dtla_utm <- st_transform(dtla, 26911)

data <- data %>%
  mutate(dist_dtla_km = as.numeric(st_distance(tracts_utm, dtla_utm)) / 1000)
```

```{r}
# Categorize the median income variable into tertiles
data <- data %>%
  mutate(
    median_income_tertile = as.character(ntile(median_income, 3)),
    median_income_quartile = as.character(ntile(median_income, 4)),
    median_income_quintile = as.character(ntile(median_income, 5)),
  )

data %>%
  select(SLEEP_CrudePrev, pw_exp_db, median_income_quintile) %>%
  drop_na() %>%
  ggplot(aes(x = pw_exp_db, y = SLEEP_CrudePrev, color = median_income_quintile, group = median_income_quintile)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ median_income_quintile)



```

```{r}
# center median_income
data <- data %>%
  mutate(
    median_income_c = median_income - mean(median_income, na.rm = TRUE)
  )
```

```{r}
#| echo: false
#| eval: false

# Check outliers for sleep / dB

# Sleep prevalence box plot
data %>%
  ggplot(., aes(y = SLEEP_CrudePrev)) +
  geom_boxplot() +
  theme_minimal()

data %>%
  ggplot(., aes(x = gte40db, y = SLEEP_CrudePrev, group = gte40db)) +
  geom_boxplot() +
  theme_minimal()


data %>%
  ggplot(., aes(x = gte50db, y = SLEEP_CrudePrev, group = gte50db)) +
  geom_boxplot() +
  theme_minimal()

data %>%
  ggplot(., aes(x = gte60db, y = SLEEP_CrudePrev, group = gte60db)) +
  geom_boxplot() +
  theme_minimal()


# Box plot of pw_exp_db in continuous noise
data %>%
  ggplot(., aes(y = pw_exp_db)) +
  geom_boxplot() +
  theme_minimal()

data <- data %>%
  mutate(
    Q1 = quantile(SLEEP_CrudePrev, 0.25, na.rm = TRUE),
    Q3 = quantile(SLEEP_CrudePrev, 0.75, na.rm = TRUE),
    IQR = IQR(SLEEP_CrudePrev, na.rm = TRUE), 
    lower = Q1 - 1.5 * IQR,
    upper = Q3 + 1.5 * IQR,
    is_outlier = ifelse(SLEEP_CrudePrev < lower | SLEEP_CrudePrev > upper, 1, 0)) 
    
data %>%
  pull(is_outlier) %>%
  table()

# data <- data %>% 
#   filter(is_outlier == 0)
```


## Group sample sizes and proportions per db cutoff value
```{r}
data %>%
  as.data.frame() %>%
  select(gte40db, gte50db, gte60db) %>%
  mutate(across(everything(), ~ factor(.x)) ) %>%
  tbl_summary() %>%
  as_tibble()
```



```{r}
# Calcualte descriptives on the SLEEP_crudePrev variable
data %>%
  as.data.frame() %>%
  select(SLEEP_CrudePrev) %>%
  drop_na() %>%
  tbl_summary(
    statistic = list(SLEEP_CrudePrev ~ "{median} ({p25}, {p75}); {mean} ({sd}); {min} - {max}"))
```

```{r}
# Calculate the width of the 95% CI
data <- data %>%
  mutate(
    SLEEP_Crude95CI_lo = as.numeric(substr(SLEEP_Crude95CI, 2, 5)),
    SLEEP_Crude95CI_hi = as.numeric(substr(SLEEP_Crude95CI, 8, 11)),
    SLEEP_Crude95CI_width = SLEEP_Crude95CI_hi - SLEEP_Crude95CI_lo,
    SLEEP_CrudeSE = SLEEP_Crude95CI_width / (2 * 1.96))

data %>%
  as.data.frame() %>%
  select(SLEEP_Crude95CI_width) %>%
  drop_na() %>%
  tbl_summary(
     statistic = list(SLEEP_Crude95CI_width ~ "{median} ({p25}, {p75}); {mean} ({sd}); {min} - {max}"))


data %>%
  ggplot(aes(x = total_pop, y = SLEEP_Crude95CI_width)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
# SES Vars with polygon for merging
ses_vars_w_poly <- data %>%
  st_drop_geometry() %>%
  select(GEOID, median_income, pct_homeowners, pct_college, median_value) %>%
  drop_na()

pca <- prcomp(
  ses_vars_w_poly %>% select(-GEOID),
  center = TRUE,
  scale. = TRUE
)

summary(pca)

pca$rotation

# Higher values as higher SES
ses_vars_w_poly$SES_PC1 <- -pca$x[, 1]

data <- data %>%
  left_join(ses_vars_w_poly %>% select(GEOID, SES_PC1), by = "GEOID")
```

# Models with a dichotomous predictor of noise exposure
<!-- The primary predictor of interest is gte50db -->

```{r}
vars <- c("gte40db", "gte50db", "gte60db")

# *** adding , "latitude", "longitude" makes the models error out system is 
# computationally singular in the errorsarlm model
co_vars <-c(
  "SES_PC1",
  # "total_pop",
  # "median_value", 
  # "pct_homeowners",
  # "median_income",
  # "pct_college",
  "pct_foreign_born", 
  "black_alr", 
  "native_alr", 
  "asian_alr", 
  "latino_alr",
  "median_ageE"
  # "lat_z",
  # "lon_z",
  # "dist_dtla_km"
  )

style <- "W" # Other option is "B"
```

```{r}
fit_dichotomous_model <- function(x) {
  # Create a copy of the main data set in the workspace
  mod_data <- data

  # Create a formula based on the input predictor variable
  # formula <- as.formula(str_c("SLEEP_CrudePrev ~ ", x)) # Crude/unadjusted model
  formula <- as.formula(str_c("SLEEP_CrudePrev ~ ", x, " + ", paste0(co_vars, collapse = " + "))) # Adjusted model
  # formula <- as.formula(str_c(
  #   "SLEEP_CrudePrev ~ ", x, " + median_income_c:", x, " + ", paste0(co_vars, collapse = " + "))) # Interaction models

  # print(formula)

  # Fit the OLS model
  model <- lm(formula, data = mod_data, na.action = na.exclude)

  # Get the residuals in a test data frame
  mod_data <- mod_data %>%
    mutate(residuals = residuals(model))

  # Drop the rows where values for the model are missing
  mod_data <- mod_data %>%
    filter(!is.na(residuals))

  # Plot the 2000 tracts that were included in the analysis
  centroids <- mod_data %>%
    filter(!!sym(x) == 1) %>%
    st_centroid()

  plt <-  mod_data %>%
    ggplot() +
    geom_sf(aes(fill = SLEEP_CrudePrev), color = NA) +
    geom_sf(
      data = centroids,
      shape = 21,
      size = .45,
      stroke = .25,
      fill = NA,
      color = "white") +
    scale_fill_viridis_c() +
    theme_minimal()

  # Construct a neighbors list from polygon list
  # *** produced a warning about some observations not having any neighbors 
  # because the intersection of tracts with noise AND tracts with a CDC Places
  # outcome results in some tracts without any neighbors
  nb <- spdep::poly2nb(
    mod_data, 
    queen = TRUE) # *** Try k nearest neighbors

  # Create the spatial weights from a neighbors list. The zero policy sets
  # weights vectors of zero length for regions without any neighbors
  lw <- spdep::nb2listw(
    nb, 
    style = style, 
    zero.policy = TRUE)

  # Moran's I test on the ols model
  moran_results <- spdep::moran.test(
    mod_data$residuals, #cannot have missing values
    lw 
  )

  # Spatial error model
  sem <- errorsarlm(
    formula,
    data = mod_data,
    listw = lw,
    zero.policy = TRUE,
    na.action = na.exclude
  )

  # Moran's I test on the spatial error model
  moran_results_sem <- spdep::moran.test(
    residuals(sem),
    lw
  )

  # GLM Models
  glm_model <- glm(formula, data = mod_data, family = Gamma(link = "log"), na.action = na.exclude)

  # OLS models with log(sleep prevalence) on LHS
  mod_data <- mod_data %>%
    mutate(SLEEP_CrudePrev = log(SLEEP_CrudePrev))

  log_model <- lm(formula, data = mod_data, na.action = na.exclude)

  # Collect output
  return(list(model, plt, moran_results, sem, moran_results_sem, log_model, glm_model))

}

results <- map(vars, fit_dichotomous_model)
```

```{r}
#| echo: false
#| eval: false
# output to an excel file
# need 

# Output the results from the ols and sem models
# View(
  bind_rows(
  broom::tidy(results[[1]][[1]]) %>% mutate(poi = "gte40db", type = "ols"),
  broom::tidy(results[[1]][[4]]) %>% mutate(poi = "gte40db", type = "sem"),
  broom::tidy(results[[2]][[1]]) %>% mutate(poi = "gte50db", type = "ols"),
  broom::tidy(results[[2]][[4]]) %>% mutate(poi = "gte50db", type = "sem"),
  broom::tidy(results[[3]][[1]]) %>% mutate(poi = "gte60db", type = "ols"),
  broom::tidy(results[[3]][[4]]) %>% mutate(poi = "gte60db", type = "sem")) %>%
  write_csv(., "D:\\rwjf_noise\\scripts\\model_results_dichotomous_w_pca.csv")
# )

# output the results from the models where the log of the dv is used
bind_rows(
  broom::tidy(results[[1]][[6]]) %>% mutate(poi = "gte40db", type = "ols"),
  broom::tidy(results[[2]][[6]]) %>% mutate(poi = "gte50db", type = "ols"),
  broom::tidy(results[[3]][[6]]) %>% mutate(poi = "gte60db", type = "ols")) %>%
  write_csv(., "D:\\rwjf_noise\\scripts\\model_results_dichotomous_logdv.csv")

# output the results from the glm models
# *** Need to confirm distribution gamma, poisson, etc.
bind_rows(
  broom::tidy(results[[1]][[7]]) %>% mutate(poi = "gte40db", type = "glm"),
  broom::tidy(results[[2]][[7]]) %>% mutate(poi = "gte50db", type = "glm"),
  broom::tidy(results[[3]][[7]]) %>% mutate(poi = "gte60db", type = "glm")) %>%
  write_csv(., "D:\\rwjf_noise\\scripts\\model_results_dichotomous_glm.csv")
```



## Models
## dB Greater than or equal to 40 dB
### OLS Regression output
```{r}
tbl_regression(results[[1]][[1]])
```

### Plot of Insufficient Sleep Prevalence overlaid with tracts where the average dB value is greater than or equal to 40 dB
```{r}
results[[1]][[2]]
```

### Moran's I on OLS model
```{r}
results[[1]][[3]]
```

### Spatial error model output
```{r}
summary(results[[1]][[4]])
```

### Moran's I on spatial error model
```{r}
results[[1]][[5]]
```

### log(sleep prevalence) vs glm() model
```{r}
# Compare glm() and model with log(sleep prevalence)

# output for OLS model with log(sleep prevalence)
summary(results[[1]][[6]])

# Plot residuals for OLS model with log(sleep prevalence)
plot(results[[1]][[6]], which = 1)

# Output for GLM model
summary(results[[1]][[7]])

# Plot residuals for GLM model
plot(fitted(results[[1]][[7]]),
     residuals(results[[1]][[7]], type = "deviance"))
abline(h = 0, lty = 2)
```

## dB Greater than or equal to 50 dB
<!-- This is one of the primary analyses -->
### OLS Regression output
```{r}
tbl_regression(results[[2]][[1]])
```

### Plot of Insufficient Sleep Prevalence overlaid with tracts where the average dB value is greater than or equal to 50 dB
```{r}
results[[2]][[2]]
```

### Moran's I on OLS model
```{r}
results[[2]][[3]]
```

### Spatial error model output
```{r}
summary(results[[2]][[4]])
```

### Moran's I on spatial error model
```{r}
results[[2]][[5]]
```

## dB Greater than or equal to 60 dB
### OLS Regression output
```{r}
tbl_regression(results[[3]][[1]])
```

### Plot of Insufficient Sleep Prevalence overlaid with tracts where the average dB value is greater than or equal to 60 dB
```{r}
results[[3]][[2]]
```

### Moran's I on OLS model
```{r}
results[[3]][[3]]
```

### Spatial error model output
```{r}
summary(results[[3]][[4]])
```

### Moran's I on spatial error model
```{r}
results[[3]][[5]]
```

# Models with binned noise predictors of noise
```{r}
data <- data %>%
  mutate(
    binned_5db = cut(
      pw_exp_db,
      breaks = seq(30, 75, by = 5),
      right = FALSE,
      include.lowest = TRUE),
    binned_10db = cut(
      pw_exp_db,
      breaks = seq(30, 80, by = 10),
      right = FALSE,
      include.lowest = TRUE)
  )
```

```{r}
# Check the distribution of the binned 5db increments
data %>%
  data.frame() %>%
  group_by(binned_5db) %>%
  count()
```

```{r}
# Check the distribution of the binned 10db increments
data %>%
  data.frame() %>%
  group_by(binned_10db) %>%
  count()
```

```{r}
# Shuffle some of the data points to avoid sparse groupings
data <- data %>%
  mutate(
    binned_5db = if_else(binned_5db %in% c("[65,70)", "[70,75]"), "[60,65)", binned_5db),
    binned_10db = if_else(binned_10db %in% c("[70,80]"), "[60,70)", binned_10db)
  )
```

```{r}
# Set the binned variables to loop over
vars <- c("binned_5db", "binned_10db")

fit_binned_model <- function(x) {
  # Create a copy of the main data set in the workspace
  mod_data <- data

  # Create a formula based on the input predictor variable
  formula <- as.formula(str_c("SLEEP_CrudePrev ~ ", x, " + ", paste0(co_vars, collapse = " + ")))
  # print(formula)

  # Fit the OLS model
  model <- lm(formula, data = mod_data, na.action = na.exclude)

  # Get the residuals in a test data frame
  mod_data <- mod_data %>%
    mutate(residuals = residuals(model))

  # Drop the rows where values for the model are missing
  mod_data <- mod_data %>%
    filter(!is.na(residuals))

  # Plot the 2000 tracts that were included in the analysis
  centroids <- mod_data %>%
    filter(!!sym(x) == 1) %>%
    st_centroid()

  plt <-  mod_data %>%
    ggplot() +
    geom_sf(aes(fill = SLEEP_CrudePrev), color = NA) +
    geom_sf(
      data = centroids,
      shape = 21,
      size = .45,
      stroke = .25,
      fill = NA,
      color = "white") +
    scale_fill_viridis_c() +
    theme_minimal()

  # Construct a neighbors list from polygon list
  # *** produced a warning about some observations not having any neighbors 
  # because the intersection of tracts with noise AND tracts with a CDC Places
  # outcome results in some tracts without any neighbors
  nb <- spdep::poly2nb(
    mod_data, 
    queen = TRUE)

  # Create the spatial weights from a neighbors list. The zero policy sets
  # weights vectors of zero length for regions without any neighbors
  lw <- spdep::nb2listw(
    nb, 
    style = style, 
    zero.policy = TRUE)

  # Moran's I test on the ols model
  moran_results <- spdep::moran.test(
    mod_data$residuals, #cannot have missing values
    lw 
  )

  # Spatial error model
  sem <- errorsarlm(
    formula,
    data = mod_data,
    listw = lw,
    zero.policy = TRUE,
    na.action = na.exclude
  )

  # Moran's I test on the spatial error model
  moran_results_sem <- spdep::moran.test(
    residuals(sem),
    lw
  )

  # GLM Models
  # glm_model <- glm(formula, data = mod_data, family = Gamma(link = "log"), na.action = na.exclude)

  # OLS models with log(sleep prevalence) on LHS
  # mod_data <- mod_data %>%
  #   mutate(SLEEP_CrudePrev = log(SLEEP_CrudePrev))

  # log_model <- lm(formula, data = mod_data, na.action = na.exclude)

  # Collect output
  return(list(model, plt, moran_results, sem, moran_results_sem))

}

results_binned <- map(vars, fit_binned_model)
```


```{r}
#| echo: false
#| eval: false

# Output to .csv files
bind_rows(
  broom::tidy(results_binned[[1]][[1]]) %>% mutate(poi = "bin_5db", type = "ols"),
  broom::tidy(results_binned[[1]][[4]]) %>% mutate(poi = "bin_5db", type = "sem"),
  broom::tidy(results_binned[[2]][[1]]) %>% mutate(poi = "bin_10db", type = "ols"),
  broom::tidy(results_binned[[2]][[4]]) %>% mutate(poi = "bin_10db", type = "sem")) %>%
write_csv(., "D:\\rwjf_noise\\scripts\\model_results_binned.csv")

```


```{r}
#| echo: false
#| eval: false


# binned 5db
summary(results_binned[[1]][[1]])


# Fit a model without the noise predictor
formula <- as.formula(str_c("SLEEP_CrudePrev ~ ", paste0(co_vars, collapse = " + ")))
no_noise <- lm(formula, data = data)
  
# Overall F test is significant, suggesting model with noise predictor results
# in a better fitting model
anova(no_noise, results_binned[[1]][[1]])

# Show the emmeans
library(emmeans)

emm <- emmeans(results_binned[[1]][[1]], ~ binned_5db)

pairs(emm)
```



# Models with a continuous predictor of noise exposure
```{r}
# Create a variable for a piece wise linear model

# Set the segment knot value
segment_knot_val <- 50

data <- data %>%
  mutate(
    segment = ifelse(pw_exp_db > segment_knot_val, pw_exp_db - segment_knot_val, 0)
  )

# Set a series of formula to iterate over
formulas <- c(
  str_c("SLEEP_CrudePrev ~ pw_exp_db + ", paste0(co_vars, collapse = " + ")),
  str_c("SLEEP_CrudePrev ~ pw_exp_db + segment + ", paste0(co_vars, collapse = " + "))
  # str_c("SLEEP_CrudePrev ~ pw_exp_db"),
  # str_c("SLEEP_CrudePrev ~ pw_exp_db + segment")

)


fit_continuous_model <- function(x) {
  # Create a copy of the main data set in the workspace
  mod_data <- data

  # Create a formula based on the input predictor variable
  formula <- as.formula(x)
  # print(formula)

  # Fit the ols model
  model <- lm(formula, data = mod_data, na.action = na.exclude)

  # Get the residuals in a test data frame
  mod_data <- mod_data %>%
    mutate(residuals = residuals(model))

  # Drop the rows where values for the model are missing
  mod_data <- mod_data %>%
    filter(!is.na(residuals))

  # Construct a neighbors list from polygon list
  # *** produced a warning about some observations not having any neighbors 
  # because the intersection of tracts with noise AND tracts with a CDC Places
  # outcome results in some tracts without any neighbors

  # Commented out to try using a knn algorithm
  nb <- spdep::poly2nb(
    mod_data, 
    queen = TRUE)

  # Alternate form of constructing the neighbors list.
  # Plot the 2000 tracts that were included in the analysis
  centroids <- mod_data %>%
    st_centroid()

  coords <- st_coordinates(centroids)
  knn <- spdep::knearneigh(coords, k = 4)
  # nb <- spdep::knn2nb(knn)

  # Create the spatial weights from a neighbors list. The zero policy sets
  # weights vectors of zero length for regions without any neighbors
  lw <- spdep::nb2listw(
    nb, 
    style = style, 
    zero.policy = TRUE)

  # Moran's I test revealed p less than .05 for all dichotomous variables
  # There is evidence for spatial auto correlation, need a spatial model
  moran_results <- spdep::moran.test(
    mod_data$residuals, #cannot have missing values
    lw 
  )

  # print(moran_results)

  # Spatial error model
  sem <- errorsarlm(
    formula,
    data = mod_data,
    listw = lw,
    zero.policy = TRUE,
    na.action = na.exclude
  )

  moran_results_sem <- spdep::moran.test(
    residuals(sem),
    lw
  )

  return(list(model, moran_results, sem, moran_results_sem))

}

results_continuous <- map(formulas, fit_continuous_model)
```

```{r}
#| echo: false
#| eval: false

# Output to .csv files
bind_rows(
  broom::tidy(results_continuous[[1]][[1]]) %>% mutate(poi = "con_db", type = "ols"),
  broom::tidy(results_continuous[[1]][[3]]) %>% mutate(poi = "con_db", type = "sem"),
  broom::tidy(results_continuous[[2]][[1]]) %>% mutate(poi = "con_db", type = "ols"),
  broom::tidy(results_continuous[[2]][[3]]) %>% mutate(poi = "con_db", type = "sem")) %>%
arrange(type) %>%
write_csv(., "D:\\rwjf_noise\\scripts\\model_results_continuous_w_pca.csv")
```

## Continuous Model 1
```{r}
# OLS model
tbl_regression(results_continuous[[1]][[1]])
```

```{r}
# Morans I on OLS Model
results_continuous[[1]][[2]]
```

```{r}
# SEM
summary(results_continuous[[1]][[3]])
```

```{r}
# Morans I on SEM
results_continuous[[1]][[4]]
```

<!-- Piecewise Model -->
## Continuous Model 2 (piece-wise/segmented model)
```{r}
tbl_regression(results_continuous[[2]][[1]])
```

```{r}
# Morans I on OLS model
results_continuous[[2]][[2]]
```

```{r}
# SEM Piecewise
summary(results_continuous[[2]][[3]])
```

```{r}
# Morans I on the Piecewise Segmented Model
results_continuous[[2]][[4]]
```

<!-- Comparison of models -->
```{r}
#| echo: false
#| eval: false


# ANOVA reveals that the segmented model has better fit than the non-segmented model.
anova(
  results_continuous[[1]][[1]],
  results_continuous[[2]][[1]])

# Splines model
mod_spline <- lm(
  SLEEP_CrudePrev ~ ns(pw_exp_db, df = 4) + median_value + pct_homeowners + pct_foreign_born + median_income + black_alr + native_alr + asian_alr + latino_alr,
  data = data
)

summary(mod_spline)

anova(
  results_continuous[[2]][[1]],
  mod_spline
)
```


## Correlation matrix between covariates
```{r}
# Correlate median income, pct_homeowners, pct_college,
View(
  cor(
  data %>% 
    st_drop_geometry() %>% 
    select(
      pw_exp_db, 
      median_value, 
      pct_homeowners, 
      median_income, 
      pct_college, 
      pct_foreign_born,
      black_alr,
      native_alr,
      asian_alr,
      latino_alr), 
  use = "pairwise.complete.obs") %>%
data.frame()
)
```