---
params:
  by_variable: 2018

title: "RWJF Helicopter Noise - Data Processing Workflow"
subtitle: "`r params$by_variable`"

format: 
  html:
    toc: true

execute:
  warning: false
  message: false # Also hides package loading messages

embed-resources: true
---


```{r}
#| eval: false
#| echo: false
# -----------------------------------------------------------------------------
# Carlos Rodriguez, PhD. CU Anschutz Dept. of Family Medicine
# RWJF HelicopterSleep

# Data processing
# The purpose of this script is to generate a file where population weighted 
# noise values are calculated for a given year, using the UCIC noise and 
# landscan data sets.
# -----------------------------------------------------------------------------
```

```{r, warning = FALSE, message = FALSE}
#%% Libraries
  library(tidyverse)
  library(tidycensus)
  library(sf)
  library(tidyterra)
  library(terra)
  library(exactextractr)
```

<!-- //////////////////////// REPORT CONFIGURATION //////////////////////// -->
<!-- This section specifies whether or not to filter the data if the by 
variable is set to by_clinic, sets rules for displaying pvalues in tables based
on whether the by variable is set to by_site_type, and sets the test argument 
to t-test if the argument is by site_type -->

```{r}
year <- params$by_variable
noise_path_prefix <- "D:\\rwjf_noise\\data\\noise\\annual\\Noise_"
noise_path_suffix <- "_Noise.csv"
noise_file <- str_c(noise_path_prefix, year, noise_path_suffix)


ls_path_prefix <- "D:\\rwjf_noise\\data\\landscan\\landscan-mosaic_timeseries-"
ls_path_suffix <- "-assets\\landscan_mosaic_timeseries_"

landscan_path <- str_c(ls_path_prefix, year, ls_path_suffix, year, ".tif")
```

# Load Landscan data
```{r}
#%% Load the landscan population raster stored locally
# has resolution of 1/1200 degree
# coordinate refeference system (crs) is 4326
pop <- rast(landscan_path)

# Older version used this file
# pop <- rast("D:\\rwjf_noise\\data\\landscan\\landscan-mosaic-unitedstates-v1-assets\\landscan-mosaic-unitedstates-v1.tif")

# pop <- rast("D:\\rwjf_noise\\data\\landscan\\landscan-global-2018-assets\\landscan-global-2018.tif")

# crs(pop)
# res(pop)
```

# Load UCIC noise data
```{r}
#%% Load and prepare noise data 
# File is .csv format with a noise level for each lat and long coordinates.
# Rename coordinate columns for convenience.
exposure_df <-read_csv(noise_file, show_col_types = FALSE) %>%
    rename(lon = Longitude, lat = Latitude) %>%
    mutate(L_night_ge32 = ifelse(L_night >= 32, 1, 0))
```

## Display data structure
```{r}
str(exposure_df)
```

## Display a sample of rows
```{r}
# Display noise exposure data frame
exposure_df %>% 
  slice_sample(n = 6)
```


# Load LA County tracts
```{r}
#%% Prepare LA County tracts object
options(tigris_progress_bar = FALSE)
options(tigris_use_cache = TRUE)

suppressMessages({
# Load LA County tracts as simple features (sf) from tidycensus
# crs loads as 4269
tracts <- get_acs(
  geography = "tract",

  # Corresponds to total population variable, not used but argument needs to be
  # set for the function to work
  variables = "B01001_001E",   
  state = "CA",
  county = "Los Angeles",
  geometry = TRUE,
  year = year
)
})

# crs(tracts)
```

## Plot raw noise data overlaid with LA county census tracts
### Create a simple features object from raw noise data
```{r}
# Set seed for sample_frac()
set.seed(12345)

# Samples a fraction of the exposure data frame for plotting, otherwise it's too
# computationally intensive to plot. Complete data set is over 2million rows.
exposure_sf <- st_as_sf(
  exposure_df %>% 
    sample_frac(.1), 
  coords = c("lon", "lat"), 
  crs = 4326)

# Transform tracts crs from 4269 to 4326
tracts <- st_transform(tracts, 4326)

# Check if both exposure and tracts have the same crs
st_crs(exposure_sf) == st_crs(tracts)
```

### Create a clipping polygon to crop maps to excude San Celemente and Santa Catalina Islands
```{r}
# Get the minimum latitude value from the exposure data to use as 
# a cutoff for removing the islands
lat_cutoff <- min(exposure_df$lat)

# Create a bb object of LA County Tracts
bb <- st_bbox(tracts)

# Modify the bb object's y minimum value
# uses the previously defined cut off value above to raise the souther edge
bb["ymin"] <- lat_cutoff   

# Create an sfc object from the modified bb object that will be used to clip
# away the islands
clip_poly <- st_as_sfc(bb)
```

### Crop tracts
```{r}
# Crop the tracts by taking the overlap between the tracts sf object
# and the clipping polygon.
la_tracts_cropped <- st_intersection(tracts, clip_poly)
```

### Plot raw noise data
- white specks appear due to random sampling of dat points
```{r}
# Plot the data using the sf objects
ggplot() +
  geom_sf(data = exposure_sf, aes(color = L_night), size = 0.5, alpha = 0.8) +
  geom_sf(data = la_tracts_cropped, fill = NA, color = "white", alpha = 0.8, linewidth = .1) +
  coord_sf() +
  labs(
    title = str_c("Average", year, " L_night values (dB) in LA County (not in Landscan space)"),
    color = "Ldn (dB)") +
  theme_minimal()
```

# Prepare LA County tracts
Workflow here is to go from from sf --> spatial vector
```{r}
# remove rows where geometry is empty
# Tracts where GEOID end in 9901:9903 with no geometry are removed
# Empty tracts must be removed to successfully perform exact_extract() in a
# downstream step, otherwise "fail to get geometry extent" errors arise.
# la_tracts_cropped <- la_tracts_cropped %>%
#   filter(!st_is_empty(la_tracts_cropped))

# Convert to spatial vector to be used for cropping
la_vect <- vect(la_tracts_cropped)

# Check if the tracts and pop have the same crs
same.crs(la_vect, pop)
```

# Crop Landscan to pixels within the LA county
```{r}
#%% Crop the landscan population raster
# crop and mask have different functions, crop is to establish the bounding box
# of pop to that of la_vect, and mask is to set any value outside the tracts, 
# but within the bbox to NA
la_pop <- mask(crop(pop, la_vect), la_vect)

cells_w_missing_values <- global(la_pop, "isNA", na.rm = FALSE)
```

```{r}
ggplot() +
  geom_spatraster(data = la_pop) +
  geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
  labs(
    title = "Landscan pop values",
    fill = "") +
  theme_minimal()
```

# Updated workflow
Since the exposure data is already in a regularly spaced grid, one approach is to rasterize the noise exposure data to its native grid after importing from .csv file and then resampling to the landscan grid.

## Rasterize exposure data frame
* Perform a linear transformation of the dB values before rasterizing
```{r}
# Linearize the dB values
exposure_df <- exposure_df %>%
  mutate(L_night_lin = 10 ^ (L_night/10))

# Create raster directly from points, instead of points to spatial vector to raster
exposure_rast <- rast(
  exposure_df[, c("lon", "lat", "L_night_lin")],
  type = "xyz",
  crs = "EPSG:4326"
)
```

```{r}
# Resolution of exposure raster is approximately 111.1 meeters by 66.7 meters.
res(exposure_rast)
```

## Resample the exposure raster
* Resample the linear transformed variables
```{r}
# Resample the exposure raster to landscan grid
exp_landscan_lin <- resample(
  exposure_rast,
  la_pop,
  method = "bilinear"
)
```

## Transform the resampled exposure raster back to dB
```{r}
# Convert exposure raster back to dB
exp_landscan_db <- 10 * log10(exp_landscan_lin)
```

## Check resolution of rasters
```{r}
# Check the resolutions of exp_landscan and la_pop
res(exp_landscan_lin)
res(exp_landscan_db)
res(la_pop)
```

```{r}
#| eval: false 
#| echo: false


## Plot the cropped exposure raster overlaid with tracts

ggplot() +
  geom_spatraster(data = exp_landscan_db) +
  geom_sf(data = la_tracts_cropped, fill = NA, color = "white", linewidth = 0.1) +
  labs(
    title = "Average L_night values (dB) in LA County (resampled to Landscan space)",
    fill = "dB") +
  theme_minimal()

ggplot() +
  geom_spatraster(data = la_pop)

```

# Approach 1: Calculate the tract level population-weighted values of noise exposure using the linear transformed dB values
```{r}
# Compare la_pop and exp_landscan geometries
# At this point la_pop is cropped to exclude Santa Catalina and San Celemente islands
compareGeom(la_pop, exp_landscan_lin)
```

```{r}
# Create a weighted value for each pixel by element-wise multiplication of the
# population and exposure rasters
weighted <- exp_landscan_lin * la_pop
```

```{r}
# Check the range of values
minmax(weighted)
```

```{r, message = FALSE}
#%% Calculated a weighted exposure variable & plot
num <- exact_extract(weighted, la_tracts_cropped, 'sum', progress = FALSE)

den <- exact_extract(la_pop, la_tracts_cropped, 'sum', progress = FALSE)

# Set the population weighted exposure variable by dividing num by den
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(pw_exp_lin = num / den)

# Convert the linearized noise levels back to dB
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(pw_exp_db = 10 * log10(pw_exp_lin))
```

## Plot population weighted exposure (dB) by tract
* Approach 1 = noise --> linear transform --> resample --> aggregate --> transform to dB
```{r}
# Plot population weighted exposure by tract
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = pw_exp_db)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```

# Create population weighted exposure variable in dB (rather than linearly transformed values, pw_exp_)
```{r}
weighted <- exp_landscan_db * la_pop
```

```{r}
minmax(weighted)
```

```{r}
#%% Calculated a weighted exposure variable & plot
num <- exact_extract(weighted, la_tracts_cropped, 'sum', progress = FALSE)
den <- exact_extract(la_pop, la_tracts_cropped, 'sum', progress = FALSE)

# Set the population weighted exposure variable by dividing num by den
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(pw_exp_db2 = num / den)

# Create a variable representing the sum of the landscan population values for each tract
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(population_sum = den)
```

## Plot differences between two approaches
```{r}
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(diff = pw_exp_db2 - pw_exp_db)
```

```{r}
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = diff)) +
    scale_fill_viridis_c() +
    labs(title = "Differences in population weighted exposure values (db2 - db)") +
    theme_minimal()
```

# Dichotomize tract level population weighted dB values at 40, 50, and 60 dB
```{r}
la_tracts_cropped <- la_tracts_cropped %>%
  mutate(
    gte40db = ifelse(pw_exp_db >= 40, 1, 0),
    gte50db = ifelse(pw_exp_db >= 50, 1, 0),
    gte60db = ifelse(pw_exp_db >= 60, 1, 0)
    )
```

## Tracts where population weighted average annual exposure is >= 40 dB
```{r}
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = gte40db)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```

```{r}
# Proportion of tracts where population weighted average annual exposure greater than or equal to 40 dB
# results in 93% of tracts out of 2493
la_tracts_cropped %>%
  pull(gte40db) %>%
  mean(., na.rm = TRUE)
```

```{r}
# Number of tracts where population weighted average annual exposure greater than or equal to 40 dB
# results in 2321 tracts
la_tracts_cropped %>%
  pull(gte40db) %>%
  sum(., na.rm = TRUE)
```

## Tracts where population weighted average annual exposure is >= 50 dB
```{r}
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = gte50db)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```

```{r}
# Proportion of tracts where population weighted average annual exposure greater than or equal to 50 dB
# results in 68% of tracts out of 2493
la_tracts_cropped %>%
  pull(gte50db) %>%
  mean(., na.rm = TRUE)
```

```{r}
# Number of tracts where population weighted average annual exposure greater than or equal to 50 dB
# results in 1695 tracts
la_tracts_cropped %>%
  pull(gte50db) %>%
  sum(., na.rm = TRUE)
```

## Tracts where population weighted average annual exposure is >= 60 dB
```{r}
la_tracts_cropped %>%
  ggplot() +
    geom_sf(aes(fill = gte60db)) +
    scale_fill_viridis_c() +
    labs(title = "Population-Weighted Exposure by Tract (dB)") +
    theme_minimal()
```

```{r}
# Proportion of tracts where population weighted average annual exposure greater than or equal to 60 dB
# results in 5.7% of tracts out of 2493
la_tracts_cropped %>%
  pull(gte60db) %>%
  mean(., na.rm = TRUE)
```

```{r}
# Number of tracts where population weighted average annual exposure greater than or equal to 60 dB
# results in 142 tracts
la_tracts_cropped %>%
  pull(gte60db) %>%
  sum(., na.rm = TRUE)
```

# Export tracts data
```{r}
# Output file to .rds
saveRDS(la_tracts_cropped, str_c("D:\\rwjf_noise\\data\\noise\\lac_pw_db_", year, ".rds"))
```